---
title: "Airborne Risk Modelling Workshop"
author: "Marco-Felipe King"
# date: '2022-11-01'
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("pacman")
pacman::p_load(shiny, deSolve,dplyr,tidyr,ggplot2,ggpubr,plotly)

```
The following document will help you work through some of the concepts we covered earlier about controlling infection and mitigation strategies. Often there are trade-offs or diminishing returns on a strategy. 

## Learning Objectives:

*By the end of the session you will be able to:*

- Evaluate the risk of airborne infection for a variety of microorganisms - Wells-Riley Model
- Evaluate the relative importance of increasing ventilation, masking or early testing on the number of infections at any one time - SEIR model.



## 1. Wells-Riley Model

The Wells-Riley equation is a method of exploring the effect of ventilation on infection risk. Recall that the equation for the steady-state model is as follows:

$$N=S\Big(1-\exp(-\dfrac{Ipqt}{Q})\Big)$$,

where $S$ is the number of infectious people, $p$ is the pulmonary breathing rate (m$^3$/h), $q$ is the quanta rate per hour, $I$ is the number of infectious people, $t$ is the time spent in the room (hours) and $Q$ is the ventilation rate in (m$^3$/h).

Use the sliders to explore the effect of each parameter

```{r Wells-Riley Sliders, echo=FALSE, warning=FALSE}
sliderInput(inputId = "quanta",label = "Quanta per hour:",
            min = 1, max = 200,value=50,step=5)

numericInput("p", "Pulmonary breathing rate (m3/h):", 1)

sliderInput(inputId = "Q",label = "Fresh air flow rate [m3/h]:",
            min = 100, max = 10000,value=200,step=200)

sliderInput("S", label = "Number of Susceptibles (-):",
            min = 30, max = 300, value = 100, step = 10)
# 
sliderInput("I", label = "Number of Infectors (-):",
            min = 1, max = 3, value = 1, step = 1)

sliderInput(inputId = "tmax",label = "Time to plot [hrs]:",
            min = 1, max = 100,value=100,step=5)

```

```{r Wells-Riley, echo=FALSE, warning=FALSE}

N <- function(p,q,Q,t,S,I){
  return(S*(1-exp(-p*I*q/Q*t)))
}

df <- reactive({
  return(data.frame(
  t=seq(0,input$tmax,by=0.1),
  value=N(input$p,input$quanta,input$Q,seq(0,input$tmax,by=0.1),input$S,input$I)
  # q250_1=N(250,1000,seq(0,8,by=0.1),input$S),
  # q1_6=N(1,1000,seq(0,8,by=0.1),input$S),
  # q250_6=N(250,6000,seq(0,8,by=0.1),input$S)
  ))
})


renderPlotly({
ggplotly(
df() %>% 
  # rename("quanta=1,ach=1"="q1_1",
  #        "quanta=250,ach=1"="q250_1",
  #        "quanta=1,ach=6"="q1_6",
  #        "quanta=250,ach=6"="q250_6",) %>% 
  # pivot_longer(!t) %>% 
  ggplot()+
  # geom_line(aes(x=t,y=value,colour=name))+
  geom_line(aes(x=t,y=value))+
  xlab("Time (hours)")+
  ylab("Number of infected")+
  scale_color_brewer(palette = "Set1")+
  # facet_wrap(~name,nrow=4)+
  hrbrthemes::theme_ipsum()+
  theme(legend.position = "")  
)
})


```

## 2. SEIR model